---
title: "Using the FGCZ Quarto report templates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the FGCZ Quarto report templates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(fgczquartotemplate)
```

## What this package provides

`fgczquartotemplate` ships one shared set of Quarto report assets so that FGCZ
analysis packages (`ezRun`, `prolfqua`, ...) all produce reports with the same
look and feel. A theme change becomes a single package bump instead of an edit
in every downstream package.

**See the layout live:**
<https://prolfqua.github.io/fgczquartotemplate/vignettes/example-report.html> — a
full report rendered with the extension (tabsets, figure + callout rows, nesting,
lightbox). The documentation site is built with
[altdoc](https://altdoc.etiennebacher.com/), which renders that report through
Quarto, so the FGCZ layout and its panel-tabsets appear exactly as they do
locally (pkgdown, by contrast, mangles Quarto tabsets). This article explains how
to *use* the templates.

**Author reports consistently:** read the
[FGCZ Quarto reports authoring guide](https://prolfqua.github.io/fgczquartotemplate/fgcz-quarto-reports.html),
which altdoc publishes directly from the
[skill source file](https://github.com/prolfqua/fgczquartotemplate/blob/main/skills/fgcz-quarto-reports/SKILL.md).
It defines the required Overview and Session Info structure, tab-depth rules,
figure layout, caption guidance, and provenance conventions.

## Two ways to use it

**Way 1 — Quarto extension** (plain `quarto render`, no R). Once per project:

```sh
quarto add prolfqua/fgczquartotemplate
```

then in the report header:

````yaml
---
title: "My report"
format: fgczquartotemplate-html
---
````

**Way 2 — R helper** (stage the files from R, no `format:` line needed) — the
subject of the rest of this article. Both routes produce the same report and
can coexist in one repo.

The assets live under `inst/quarto/`:

```{r list-assets}
list.files(fgcz_quarto_dir())
```

| File | Role |
|------|------|
| `_metadata.yml` | Shared format defaults, applied automatically to every `.qmd` in the directory. |
| `fgcz.scss` | Theme overrides — tabset/card styling, figure rows. |
| `fgcz_header_quarto.html` | FGCZ header, injected via `include-in-header`. |
| `fgcz-plot-finder.html` | Opt-in Find and Download toolbar. |
| `fgcz-buttons.lua` | Validates and applies toolbar selections for the Quarto extension. |
| `fgcz-report-overview.svg` | Visual abstract copied with the starter report. |
| `template.qmd` | Generic starter report demonstrating the layout patterns. |

## Why the assets must sit next to the report

Quarto applies a file named `_metadata.yml` automatically to every `.qmd` in
its directory (and subdirectories) — that is how the styling attaches without
any front-matter reference. The file in turn names `fgcz.scss` and
`fgcz_header_quarto.html` by **bare filename**:

```{r peek-yml}
writeLines(head(readLines(fgcz_quarto_dir("_metadata.yml")), 30))
```

Quarto resolves those names relative to the directory of the **input `.qmd`**,
not relative to the YAML that mentions them. So all three files must physically
sit next to the report at render time. That is exactly what
`fgcz_copy_assets()` and `fgcz_render()` take care of.

## The portable report header

Because `_metadata.yml` attaches by directory, a report carries **no**
FGCZ-specific front matter at all — just its own params and title:

````yaml
---
params:
  reportTitle: "CountQC"
title: "`r knitr::inline_expr("params$reportTitle")`"
---
````

Rendered with a bare `quarto render CountQC.qmd`, it still picks up the full
styling — the package need not even be installed, as long as the three assets
are in the directory.

## Staging the assets

`fgcz_copy_assets()` copies the three styling files into a directory. Here we
stage them into a temporary directory and confirm they landed:

```{r copy-assets}
dir <- file.path(tempdir(), "demo-report")
dir.create(dir, showWarnings = FALSE)

staged <- fgcz_copy_assets(dir)
basename(staged)
file.exists(staged)
```

## Bootstrapping a new report

`fgcz_use_template()` copies the starter `template.qmd` into a directory
together with the styling assets it needs, so you have a runnable report in one
call:

```{r use-template}
qmd <- fgcz_use_template(file.path(tempdir(), "new-report"),
                         to = "my_report.qmd",
                         overwrite = TRUE)
list.files(dirname(qmd))
```

## Rendering

`fgcz_render()` stages the assets and then calls `quarto::quarto_render()`. It
needs the Quarto CLI, so the call below is shown but not run in this vignette:

```{r render, eval = FALSE}
fgcz_render(qmd, execute_params = list(reportTitle = "My analysis"))
fgcz_render(qmd, buttons = TRUE) # both toolbar buttons; FALSE remains the default
fgcz_render(qmd, buttons = "search") # Find only; "download" selects Download
```

That is the single call downstream packages use in place of a bare
`quarto::quarto_render()`.

With the Quarto extension route, opt into the toolbar with
`include-after-body` and select controls using a top-level `fgcz-buttons:` key:

````yaml
format:
  fgczquartotemplate-html:
    include-after-body: _extensions/fgczquartotemplate/fgcz-plot-finder.html
fgcz-buttons: [search, download]
````

Use a single name for one button. Omitting `fgcz-buttons` shows both controls;
an unknown name stops the render with a validation error.

## Using it from a downstream package

1. Add `fgczquartotemplate` to `Imports` (and, since it is distributed on
   GitHub, `Remotes: prolfqua/fgczquartotemplate`).
2. Replace `quarto::quarto_render()` calls with
   `fgczquartotemplate::fgcz_render()`.
3. Leave report front matter free of any styling reference — the staged
   `_metadata.yml` attaches automatically.
4. Git-ignore the staged copies (`_metadata.yml`, `fgcz.scss`,
   `fgcz_header_quarto.html`, and `fgcz-plot-finder.html`) in the render
   directory so the packaged assets remain the single source of truth.
