---
title: "Using the FGCZ Quarto report templates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the FGCZ Quarto report templates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(fgczquartotemplate)
```

## What this package provides

`fgczquartotemplate` ships one shared set of Quarto report assets so that FGCZ
analysis packages (`ezRun`, `prolfqua`, ...) all produce reports with the same
look and feel. A theme change becomes a single package bump instead of an edit
in every downstream package.

The assets live under `inst/quarto/`:

```{r list-assets}
list.files(fgcz_quarto_dir())
```

| File | Role |
|------|------|
| `_fgcz-report.yml` | Shared format defaults, pulled into a report via `metadata-files`. |
| `fgcz.scss` | Theme overrides — tabset/card styling, figure rows. |
| `fgcz_header_quarto.html` | FGCZ header, injected via `include-in-header`. |
| `template.qmd` | Generic starter report demonstrating the layout patterns. |

## Why the assets must sit next to the report

The three styling files reference each other by **bare filename** — the format
YAML names `fgcz.scss` and `fgcz_header_quarto.html`:

```{r peek-yml}
writeLines(head(readLines(fgcz_quarto_dir("_fgcz-report.yml")), 30))
```

Quarto resolves those names relative to the directory of the **input `.qmd`**,
not relative to the YAML that mentions them. So all three files must physically
sit next to the report at render time. That is exactly what
`fgcz_copy_assets()` and `fgcz_render()` take care of.

## The portable report header

A report never hard-codes a path into the package. It keeps the same header it
would have if the assets were local:

````yaml
---
params:
  reportTitle: "CountQC"
title: "`r knitr::inline_expr("params$reportTitle")`"
metadata-files: ["_fgcz-report.yml"]
---
````

## Staging the assets

`fgcz_copy_assets()` copies the three styling files into a directory. Here we
stage them into a temporary directory and confirm they landed:

```{r copy-assets}
dir <- file.path(tempdir(), "demo-report")
dir.create(dir, showWarnings = FALSE)

staged <- fgcz_copy_assets(dir)
basename(staged)
file.exists(staged)
```

## Bootstrapping a new report

`fgcz_use_template()` copies the starter `template.qmd` into a directory
together with the styling assets it needs, so you have a runnable report in one
call:

```{r use-template}
qmd <- fgcz_use_template(file.path(tempdir(), "new-report"),
                         to = "my_report.qmd",
                         overwrite = TRUE)
list.files(dirname(qmd))
```

## Rendering

`fgcz_render()` stages the assets and then calls `quarto::quarto_render()`. It
needs the Quarto CLI, so the call below is shown but not run in this vignette:

```{r render, eval = FALSE}
fgcz_render(qmd, execute_params = list(reportTitle = "My analysis"))
```

That is the single call downstream packages use in place of a bare
`quarto::quarto_render()`.

## Using it from a downstream package

1. Add `fgczquartotemplate` to `Imports` (and, since it is distributed on
   GitHub, `Remotes: prolfqua/fgczquartotemplate`).
2. Replace `quarto::quarto_render()` calls with
   `fgczquartotemplate::fgcz_render()`.
3. Keep report headers as `metadata-files: ["_fgcz-report.yml"]`.
4. Git-ignore the staged copies in the render directory so the packaged assets
   remain the single source of truth.
